---
title: 'Forecasting energy consumption: exploratory data analysis'
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---

As dataset is large, I did some initial data exploration to demonstrate proof of concept using the first 1 million rows.


```{r}
library(data.table)
library(dplyr)
library(prophet)
```

### Read in and prepare data
```{r}
# read in first csv file of 1 million rows
csv_name = 'C:/Users/User/Documents/my_code_files/Exercises/Forecasting London energy consumption Oct 2017/smart_meter_data/separate_csvs/Power-Networks-LCL-June2015(withAcornGps)v2_2.csv'
mydata <- fread(csv_name, drop = c("Acorn", "Acorn_grouped"))
head(mydata)

# Compute daily consumption per household  
daily_kwh_per_household <- mydata %>%  
  # use households on standard tariff only 
  filter(stdorToU=="Std") %>%
  # rename variables for ease of reference
  rename(kwh = 'KWH/hh (per half hour)') %>%
  # extract date from datetime
  mutate(
    dt = as.POSIXct(paste(DateTime)),
    day = as.Date(strftime(dt, format = "%D"), "%m/%d/%y")) %>%
  # compute total daily consumption for each household
  group_by(day, LCLid) %>%
  summarise(total_kwh = sum(as.numeric(kwh)))
  
str(daily_kwh_per_household)
length(unique(daily_kwh_per_household$LCLid)) # 27 households

# Save as CSV
# fwrite(daily_kwh_per_household, file = "daily_kwh_per_household_csv2_2.csv")
```

### plots - by household
```{r}
ggplot(data = daily_kwh_per_household, 
mapping = aes(x = day, y = total_kwh)) +
  geom_smooth(aes(colour = LCLid), show.legend = FALSE)
```

### Get mean per day across households
```{r}
mpd <- summarise(daily_kwh_per_household, mean_kwh = mean(total_kwh))
head(mpd)

# plot means
ggplot(data = mpd, 
       mapping = aes(x = day, y = mean_kwh)) +
  geom_point(show.legend = FALSE) + 
  geom_smooth(show.legend = FALSE)
```

### Prepare df for time series analysis
```{r}
mdkwh <- select(ungroup(mpd), ds = day, y = mean_kwh)
head(mdkwh)
```

### Fit additive regression model
```{r}
arm <- prophet(mdkwh, growth = "linear",
               yearly.seasonality=TRUE, 
               weekly.seasonality=FALSE)
```

### predict consumption for the next year
```{r}
# first make list of dates to predict for
future <- make_future_dataframe(arm, periods = 365)
tail(future) # view last dates

# make the forecast
forecast <- predict(arm, future)
tail(forecast[c('ds', 'yhat', 'yhat_lower', 'yhat_upper')])
```

### simple plot with raw data plus forecast 1y into future
```{r}
ggplot(data = arm$history, mapping = aes(x = ds, y = y)) +
  geom_point() +
  geom_smooth(data = forecast, mapping = aes(x = ds, y = yhat)) + 
  geom_ribbon(data = forecast, mapping = aes(x = ds, y = yhat,
                                             ymin = yhat_lower,
                                             ymax = yhat_upper), alpha = 1/3) +
  ylab("Consumption per household (kWh)") +
  xlab("")
```

### plot seasonal components
```{r}
prophet_plot_components(arm, forecast)
```
